---
title: "drafts"
---

# Nonparametric regressions


In this part, we continue working with the corruption across the globe. However, we will use another type of the variable for the corruption. First, we extract all the data we need from the V-Dem dataset. Those include `country_name`, `year`, Political corruption index (`e_v2x_corr_5C`) and GDP per capita (`e_gdppc`) variables.


```{r}
corruption_gdp = vdem %>%
  select(country_name, year, e_v2x_corr_5C, e_gdppc) %>%
  rename(corruption = e_v2x_corr_5C,
         gdp_percap = e_gdppc) %>%
  filter(year >= 2000)
```

Explore the variable `corruption` below. What do you notice?

```{r}
#| warning: false

ggplot(corruption_gdp) +
  geom_histogram(aes(x = corruption))
```

Let's check which values this variable takes. And we want to know it's class too.

```{r}
table(corruption_gdp$corruption)
class(corruption_gdp$corruption)
```

Apparently, those are categories. Thus, we need to treat them as factors instead of numeric.

```{r}
corruption_gdp$corruption = as.factor(corruption_gdp$corruption)
```

Great! Now, let's reproduce the logic behind the Simple Local Averages regression. Visualize GDP per capita distribution across corruption index groups.

```{r}
#| warning: false

ggplot(corruption_gdp,
       aes(x = corruption,
           y = gdp_percap)) +
  geom_point() +
  labs(x = "Corruption Index (Categorical)",
       y = "GDP Per Capita")
```

Now, we should add a local average, i.e. an average in each group. Take a moment to understand the code.

```{r}
#| warning: false

ggplot(corruption_gdp,
       aes(x = corruption,
           y = gdp_percap)) +
  geom_point() +
  stat_summary(fun = mean, 
               color = "red") +
  labs(x = "Corruption Index (Categorical)",
       y = "GDP Per Capita")
```


# Legislative capacity and electoral turnout

Election turnout (A) (v2eltrnout)
Legislative constraints on the executive index (D) (v2xlg_legcon)

Are the capacities of the legislature associated with electoral turnout? Is there an association between legislative constraints on the executive and electoral turnout? 

```{r}
#| eval: false
legislative_data = vdem %>% 
  select(country_name, year, v2xlg_legcon, v2eltrnout, e_v2x_polyarchy_5C) %>%
  rename(legislative_constraints = v2xlg_legcon,
         turnout = v2eltrnout,
         regime = e_v2x_polyarchy_5C) %>%
  filter(year > 2020 & regime >= 0.5) 

legislative_data %>%
  ggplot(aes(x = legislative_constraints, y = turnout)) +
  geom_point() +
  geom_smooth(method = "lm")

lm(turnout ~ legislative_constraints, legislative_data) %>%
  summary()

library(estimatr)

```



## Fixed Effects

Let's work with data from the George Ward's article "Happiness and voting: Evidence from four decades of elections in Europe". It's available [here](https://doi.org/10.1111/ajps.12492). Let us list some variables from the [codebook](https://doi.org/10.7910/DVN/QWLGGN):

`vote_share_cab` - % votes won by cabinet parties

`satislfe_survey_mean` - Life Satisfaction

```{r}
ward_data = read.csv("data/ward2020_subset.csv")

ward_data$country = as.factor(ward_data$country)



model = lm(vote_share_cab ~ satislfe_survey_mean +
             country,
           data = ward_data)


library(ggeffects)


ggplot(ward_data, aes(x = satislfe_survey_mean, y = vote_share_cab)) +
  geom_point() +
  geom_smooth(method = "lm")


ggpredict(model, terms= c("satislfe_survey_mean", "country")) %>%
  plot(show_data = T, show_ci  = FALSE, colors = "okabe-ito", alpha = 0.5, line_size = 1) 


```


## Interactions


```{r}
whr_subset = whr %>%
  select(Country_name, Ladder_score, Social_support, Perceptions_of_corruption, Continent)

whogov_subset = whogov %>%
  filter(year == 2020) %>%
  select(country_name, n_total, n_minister, n_individuals, system_category, indep)

whr_whogov = whr_subset %>%
  left_join(whogov_subset, by = c("Country_name" = "country_name")) 

whr_whogov = whr_whogov %>%
  mutate(system_category = case_when(system_category == "Parliamentary democracy" ~ "democracy",
                                     system_category == "Presidential democracy" ~ "democracy",
                                     system_category == "Mixed democratic" ~ "democracy",
                                     is.na(system_category) ~ NA,
                                     T ~ "autocracy"))

whr_whogov$system_category = as.factor(whr_whogov$system_category)
```


```{r}
ggplot(whr_whogov, aes(x = n_individuals, y = Ladder_score)) +
  geom_point(aes(color = system_category)) +
  geom_smooth(method = "lm")

model_in = lm(Social_support ~ n_individuals * indep, whr_whogov)
summary(model_in)

ggpredict(model_in, terms= c("n_individuals", "indep")) %>%
  plot(show_ci = F, show_data = T) 

```
