---
title: "Untitled"
---

```{r}
library(tidyverse)
library(readxl)
library(estimatr)
library(modelsummary)

df = read_excel("data/cpds.xlsx")
```

outlays Total outlays (disbursements) of general government as a percentage of GDP

gov_right1 Government composition: cabinet posts of right-wing parties in percentage of total cabinet
posts. 

prefisc_gini ~ openc + servadmi_pmp + year + country

-   `prefisc_gini` - Gini index

-   `openc` - Openness of the economy

-   `servadmi_pmp` - Public and mandatory private employment services and administration as a percentage of GDP.



```{r}
df$outlays = as.numeric(df$outlays)
df$emprot_reg = as.numeric(df$emprot_reg)
df$openc = as.numeric(df$openc)

model = lm(prefisc_gini ~ openc + servadmi_pmp + country + year, df)
model_rob = lm_robust(prefisc_gini ~ openc + servadmi_pmp + country + year, df, se_type = "HC2")

summary(model)
summary(model_rob) 

modelsummary(list(model, model_rob), 
             stars = T,
             coef_omit = c("country|year"))
```


```{r}
plot(model, 1)
plot(model, 2)

library(lmtest)
bptest(model)
```


```{r}
ggplot() +
  geom_point(aes(y = model$residuals,
                 x = predict(model))) +
  geom_hline(yintercept = 0)

```


cluster: thus, we can control for the confounders on the country level