---
title: "GLMs and Quarter Review"
subtitle: "Week 9"
date: 2025-03-06
format: 
  html:
    embed-resources: true
toc: true
---


# Before we start

-   Any questions?


<center>
<a href="https://artur-baranov.github.io/nu-ps405-ds/scripts/ps405-d_9.qmd" class="btn btn-primary" role="button" download="ps405-d_9.qmd" style="width:200px" target="_blank">Download script</a>
</center>
<br>
<center>
<a href="https://artur-baranov.github.io/nu-ps405-ds/data/..." class="btn btn-primary" role="button" download="..." style="width:200px" target="_blank">Download data</a>
</center>


# Review of the previous week

::: callout-note
## Review

Load `tidyverse` library

```{r}
#| message: false
library(tidyverse)
```

Load the European Social Value data (`ess.csv`).

```{r}

```

Rename the variables: 

-   `pplfair` to `trust`

-   `trstplt` to `trust_politicians`

-   `rlgdgr` to `religion`

```{r}

```

Run the following model. DV: `happy`, IVs: `trust`, `trust_politicians` and `religion`. Present the `summary()`.

```{r}

```

Using `bptest()` from the `lmtest` library, test if the model is homoscedastic. Interpret the results of Breusch-Pagan test.

```{r}

```

Draw residuals vs fitted plot. Does it match the BP test results? You can use `plot(..., which = 1)`.

```{r}

```

Calculate the cutoff for the Cook's distance using the formula: $\frac{4}{N-k-1}$. Plot the Cook's distance using `plot()` and add a cutoff line using `abline()`. Are there any influential points?

```{r}

```

Set up a model without influential points. How different the estimates are? Present a simple `modelsummary()` comparison.

```{r}
#| eval: false

model_no = lm(happy ~ trust + trust_politicians + religion, ess[-c(...),])

library(...)
modelsummary(list(...),
             stars = T)
```

Draw a Q-Q plot using `plot(..., which = 2)` for the full model. Are the residuals distributed normally?

```{r}

```

Using `shapiro.test()` test the normality of the `residuals` in the model. Interpret the results.

```{r}

```

What else we have to check? How would you do that? No code needed but feel free to rely on the previous lab for this question.

:::


# Agenda

-   Recap of what we did

-   And a short introduction to GLM alongside

# Model Building

Today we are working with Transitional Justice Evaluation Tools dataset. Let's load the data.

```{r}
#| eval: false

tjet = read.csv("data/tjet.csv")
```

Let's check the first observations

```{r}
#| eval: false

head(tjet)
```


We are going to us the following variables:


-   `dtr`	democratic transition sample, binary

-   `tcs_operated`	truth commission operated, binary

-   `trials_domestic`	count of domestic trials per country and start year, by country of the accused



```{r}
mod = glm(dtr ~ trials_domestic + tcs_operated, tjet, family = binomial(link = "logit")) 
summary(mod)

library(ggeffects)
ggpredict(mod, term = c("trials_domestic", "tcs_operated")) %>%
  plot(show_data = T)
```


https://rstudio.github.io/cheatsheets/

What's next?