---
title: "Model Diagnostics"
subtitle: "Week 8"
date: 2025-02-27
format: 
  html:
    embed-resources: true
toc: true
---

# Before we start

-   Any questions?

-   We're almost done with the class! What you think we should recap? Feel free to leave any thoughts in this [form](https://forms.gle/uCBiJp8fB1a5xcCs7) after the class

-   To get the extra points for the exercises you will need to complete only 5 exercises. 

<center>
<a href="https://artur-baranov.github.io/nu-ps405-ds/scripts/ps405-d_8.qmd" class="btn btn-primary" role="button" download="ps405-d_8.qmd" style="width:200px" target="_blank">Download script</a>
</center>
<br>
<center>
<a href="https://artur-baranov.github.io/nu-ps405-ds/data/cpds.xlsx" class="btn btn-primary" role="button" download="cpds.xlsx" style="width:200px" target="_blank">Download data</a>
</center>

# Review of the previous week

::: callout-note
## Review

Load `tidyverse` library

```{r}
library(tidyverse)
```

Load the Comparative Political Dataset data (`cpds.xlsx`).

```{r}
library(readxl)
cpds = read_excel("data/cpds.xlsx")
```


prop Electoral system

effpar_ele Effective number of parties 


```{r}
library(lmtest)
lm(effpar_ele ~ as.factor(prop) + country + as.factor(year), cpds) %>%
  summary()
```

```{r}
library(estimatr)

lm_robust(effpar_ele ~ as.factor(prop), cpds,
          fixed_effects =  country + as.factor(year),
          se_type = "HC2")

```


:::


# Agenda

-   Continuing the diagnostics

-   Cook's distance


```{r}
cooks.distance()
```

4 / N - k (covariates) - 1
